---
layout: post
title:  "Using SmarterPoland-package in accessing Eurostat data"
date:   2013-08-06 11:00:45
categories: R-project
---


[SmarterPoland-package](http://cran.r-project.org/web/packages/SmarterPoland/index.html) provides a straghtforward connection to Eurostat data. It is uninformatively described as:

>A set of tools developed by the Foundation SmarterPoland.pl
>Tools for accessing and processing datasets presented on the blog SmarterPoland.pl.

But in real terms it has functionality only towards Eurostat. Here is a brief demo how you can search for *material deprivation* and then create a line plot at NUTS2 level.

## Search

```{r, message=FALSE, warning=FALSE}
library(SmarterPoland)
searchresults <- grepEurostatTOC("material deprivation")
df <- getEurostatRCV(kod = "ilc_mddd21")
```

## Plot at nuts2-level

```{r, message=FALSE, warning=FALSE}
# time variable into numerical
df$time <- as.numeric(levels(df$time))[df$time]
cname <- subset(df, time == 2011)

# plot
library(ggplot2)
ggplot(df, aes(x=time,y=value, color=geo,group=geo)) + 
  geom_point() + geom_line() +
  geom_text(data=cname, 
            aes(x=time,y=value,label=geo), hjust=-0.3) +
  theme(legend.position="none") +
  labs(title="material deprivation in EU at NUTS2-level",
       y="population share (%)") +
  coord_cartesian(xlim=c(2003,2012)) +
  scale_x_continuous(breaks = 2003:2011)
```


## Plot at nuts1-level

```{r, message=FALSE, warning=FALSE}
# subset geo-names only lenght of 2 characters
df$geo <- as.character(df$geo)
df$geo.n <- nchar(df$geo)
df <- subset(df, geo.n < 3)

cname <- subset(df, time == 2011)

# plot
library(ggplot2)
ggplot(df, aes(x=time,y=value, color=geo,group=geo)) + 
  geom_point() + geom_line() +
  geom_text(data=cname, 
            aes(x=time,y=value,label=geo), hjust=-0.3) +
  theme(legend.position="none") +
  labs(title="material deprivation in EU at NUTS2-level",
       y="population share (%)") +
  coord_cartesian(xlim=c(2003,2012)) +
  scale_x_continuous(breaks = 2003:2011)
```





[jekyll-gh]: https://github.com/mojombo/jekyll
[jekyll]:    http://jekyllrb.com
